DeepSphere: a graph-based spherical CNN

### 2019-11-15 main.tex
fix R(g)

\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{prop}{Proposition}
\newtheorem*{remark}{Remark}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\renewcommand{\b}[1]{{\bm{#1}}}  % bold symbol
\newcommand{\bO}{\mathcal{O}}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\S}{\mathbb{S}}
\newcommand{\G}{\mathcal{G}}  % graph
\newcommand{\V}{\mathcal{V}}  % vertex set
\newcommand{\E}{\mathcal{E}}  % edge set
\newcommand{\laplacian}{\Delta_{\mathbb S^2}}
\newcommand{\Rg}{\mathcal R (g)}
\newcommand{\seminorm}[1]{\left|#1\right|}
\newcommand{\Ln}{\hat L_n^{t_n}}

@@

\todo{
rebuttal:
\begin{itemize}
	\item \mdeff{Why two definitions of rotations, $R(g)$ and $\mathcal{R}_g$?} [Martino]
    \item answer commenter
	\item add somewhere (intro?): exploit symmetries to reduce sample cost $\Rightarrow$ we don't discuss symmetries much
	\item \citet{mudigonda2017climateevents} use Integrated Water Vapor (IWV) only. That most probably is the same as our TMQ, which is most probably the most informative feature. TODO: run ours with TMQ alone.
	\item Show the three samplings in a figure?
	\item better abstract (Friday's talk abstract?)
	\item conclusion: less speculation?
	\item check feedback from ICLR'20 LTS2 reading group
	\item \mdeff{check my Telegram notes}
\end{itemize}
}

@@

% 0.2 Explain why the Laplacian is so important...
By definition, the (spherical) graph convolution is rotation equivariant if and only if it commutes with the rotation operator defined as $R(g), g\in SO(3)$: $R(g) f(x) = f\left(g^{-1} x \right)$.
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian (\eqnref{graph_conv}).
Hence, if $R(g)$ commutes with $\b{L}$, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, $h(\b{L})$ is rotation equivariant if and only if
% 0.3 Formulate the problem
\begin{equation*} %\label{eq:equivariance}
	\b{R}_\V(g) \b{L} \b{f} = \b{L} \b{R}_\V(g) \b{f}, \hspace{1cm} \forall f\in F_\V \text{ and } \forall g\in SO(3),
\end{equation*}
where $\b{R}_\V(g) = T_\V R(g) T_\V^{-1}$.
For an empirical evaluation of equivariance, we define the \textit{normalized equivariance error} for a signal $\b{f}$ and a  rotation $g$ as
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_\V(g) \b{L} \b{f} - \b{L} \b{R}_\V(g) \b{f}} }{\norm {\b{L} \b{f}}}\right)^2.
\end{equation}
More generally for a class of signals $f \in C \subset F_\V$, the \textit{mean equivariance error} defined as
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g)
\end{equation}
represents the overall equivariance error.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

@@

% 2) explain our approach
To achieve equivariance to all rotations, we take inspiration from \citet{belkin2005towards}.
They prove that for a \emph{random uniform sampling}, the graph Laplacian $\b{L}$ built from weights
\begin{equation} \label{eqn:weights:belkin}
	w_{ij} = e^{- \frac{1}{4t} \norm{x_i - x_j}^2}
\end{equation}
converges to the Laplace-Beltrami operator $\Delta_{\mathbb{S}^2}$ as the number of samples grows to infinity.
This result is a good starting point as $\Delta_{\mathbb{S}^2}$ commutes with rotation, i.e., $\Delta_{\mathbb{S}^2}R(g) = R(g)\Delta_{\mathbb{S}^2}$.
% and (ii) is diagonalized by the spherical harmonics.
While the weighting scheme is full (i.e., every vertex is connected to every vertex), most weights are small due to the exponential.
We hence make an approximation to limit the cost of the convolution \eqnref{graph_conv} by only considering the $k$ nearest neighbors to each vertex.
%In practice, the tricky part is to set $t$ in a non-asymptotic regime.
Given $k$, the optimal kernel width $t$ is found by searching for the minimizer of \eqref{eq:mean equivariance error}.
\Figref{kernel_widths} shows the optimal kernel widths found for various resolutions of the HEALPix sampling.
As predicted by the theory, $t_n \propto n^\beta, \beta \in \R$.
Importantly however, the optimal $t$ also depends on the number of neighbors $k$.

@@

\paragraph{Theoretical convergence.}
\begin{wrapfigure}{r}{5cm}
	\centering
	\includegraphics[width=\linewidth]{patch}
	\caption{The patch around the $i$th sampling point.}
\end{wrapfigure}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling.
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
\begin{equation} \label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=1}^{n} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
\end{equation}
This operator extends the graph Laplacian operator previously defined with the weighting scheme \eqnref{weights:belkin} to each point of the the sphere.
% \mdeff{Can we avoid to introduce general manifolds and stay on the sphere?} \nati{yes}
As the radius of the kernel $t$ will be adapted to the number of samples, we also rescale the graph Laplacian operator as $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$.
Hence, we have $\b{L}_n^t \b{f} = T_\V L_n^t f$.
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling $\V$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
Define $d^{(n)} := \max_{i=1, \dots, n} d_i$ and $A^{(n)} := \max_{i=1, \dots, n} A_i$.
%Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere
%\mdeff{Already defined in problem formulation.}
%\begin{equation} \label{eq:rotation operator}
%	\mathcal R_g f(x) := f(g^{-1}(x)).
%\end{equation}
\begin{theorem}
	For a sampling $\V$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*}
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the rescaled graph Laplacian (the discrete counterpart of $\hat L^t_n$) converges pointwise towards the Laplace-Beltrami operator $\Delta_{\S^2}$ in the sense explained in \citep{belkin2007convergence}.
This is a major step towards equivariance as the Laplace-Beltrami operator does commute with rotation. Based on this property, we show that the graph Laplacian is equivariant to rotation in the limit of infinite sampling.

\begin{theorem}\label{theo:equivariance}
Under the hypothesis of theorem~\ref{theo:pointwise convergence for a regular sampling}, the rescaled graph Laplacian commutes with any rotation, in the limit of infinite sampling, i.e.,
\begin{equation*}
	\forall y\in\mathbb S^2\quad\left| \mathcal R(g) \hat L_n^{t_n} f (y) - \hat L_n^{t_n} \mathcal R(g) f(y) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{theorem}

@@

\section{Proof of Theorem~\ref{theo:equivariance}}
\begin{proof}
Fix $x\in \mathbb S^2$. Since any rotation $\mathcal R_g$ is an isometry, and the Laplacian $\Delta$ commutes with all isometries of a Riemanniann manifold, and defining $\mathcal R_g f =: f'$ for ease of notation, we can write that
	\begin{align*}
			\seminorm{\Rg \Ln f (x) - \Ln \Rg f(x) } &\leq \seminorm{\Rg \Ln f (x) - \Rg \laplacian f(x)} + \seminorm{\Rg \laplacian f (x) - \Ln \Rg f(x)} =\\
			& =\seminorm{\Rg (\Ln f - \laplacian f ) (x)} + \seminorm{ \laplacian f' (x) - \Ln f'(x)} \leq \\
			& \leq \seminorm{(\Ln f - \laplacian f ) (g^{-1}(x))} + \seminorm{ \laplacian f' (x) - \Ln f'(x)}
	\end{align*}

>>>>

\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{prop}{Proposition}
\newtheorem*{remark}{Remark}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\renewcommand{\b}[1]{{\bm{#1}}}  % bold symbol
\newcommand{\bO}{\mathcal{O}}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\S}{\mathbb{S}}
\newcommand{\G}{\mathcal{G}}  % graph
\newcommand{\V}{\mathcal{V}}  % vertex set
\newcommand{\E}{\mathcal{E}}  % edge set
\newcommand{\laplacian}{\Delta_{\mathbb S^2}}
\newcommand{\Rg}{\mathcal{R}(g)}
\newcommand{\seminorm}[1]{\left|#1\right|}
\newcommand{\Ln}{\hat L_n^{t_n}}

@@

\todo{
rebuttal:
\begin{itemize}
    \item answer commenter
	\item add somewhere (intro?): exploit symmetries to reduce sample cost $\Rightarrow$ we don't discuss symmetries much
	\item \citet{mudigonda2017climateevents} use Integrated Water Vapor (IWV) only. That most probably is the same as our TMQ, which is most probably the most informative feature. TODO: run ours with TMQ alone.
	\item Show the three samplings in a figure?
	\item better abstract (Friday's talk abstract?)
	\item conclusion: less speculation?
	\item check feedback from ICLR'20 LTS2 reading group
	\item \mdeff{check my Telegram notes}
\end{itemize}
}

@@

% 0.2 Explain why the Laplacian is so important...
By definition, the (spherical) graph convolution is rotation equivariant if and only if it commutes with the rotation operator defined as $\Rg, g\in SO(3)$: $\Rg f(x) = f\left(g^{-1} x \right)$.
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian (\eqnref{graph_conv}).
Hence, if $\Rg$ commutes with $\b{L}$, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, $h(\b{L})$ is rotation equivariant if and only if
% 0.3 Formulate the problem
\begin{equation*} %\label{eq:equivariance}
	\b{R}_\V(g) \b{L} \b{f} = \b{L} \b{R}_\V(g) \b{f}, \hspace{1cm} \forall f\in F_\V \text{ and } \forall g\in SO(3),
\end{equation*}
where $\b{R}_\V(g) = T_\V \Rg T_\V^{-1}$.
For an empirical evaluation of equivariance, we define the \textit{normalized equivariance error} for a signal $\b{f}$ and a  rotation $g$ as
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_\V(g) \b{L} \b{f} - \b{L} \b{R}_\V(g) \b{f}} }{\norm {\b{L} \b{f}}}\right)^2.
\end{equation}
More generally for a class of signals $f \in C \subset F_\V$, the \textit{mean equivariance error} defined as
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g)
\end{equation}
represents the overall equivariance error.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

@@

% 2) explain our approach
To achieve equivariance to all rotations, we take inspiration from \citet{belkin2005towards}.
They prove that for a \emph{random uniform sampling}, the graph Laplacian $\b{L}$ built from weights
\begin{equation} \label{eqn:weights:belkin}
	w_{ij} = e^{- \frac{1}{4t} \norm{x_i - x_j}^2}
\end{equation}
converges to the Laplace-Beltrami operator $\Delta_{\mathbb{S}^2}$ as the number of samples grows to infinity.
This result is a good starting point as $\Delta_{\mathbb{S}^2}$ commutes with rotation, i.e., $\Delta_{\mathbb{S}^2}\Rg = \Rg\Delta_{\mathbb{S}^2}$.
% and (ii) is diagonalized by the spherical harmonics.
While the weighting scheme is full (i.e., every vertex is connected to every vertex), most weights are small due to the exponential.
We hence make an approximation to limit the cost of the convolution \eqnref{graph_conv} by only considering the $k$ nearest neighbors to each vertex.
%In practice, the tricky part is to set $t$ in a non-asymptotic regime.
Given $k$, the optimal kernel width $t$ is found by searching for the minimizer of \eqref{eq:mean equivariance error}.
\Figref{kernel_widths} shows the optimal kernel widths found for various resolutions of the HEALPix sampling.
As predicted by the theory, $t_n \propto n^\beta, \beta \in \R$.
Importantly however, the optimal $t$ also depends on the number of neighbors $k$.

@@

\paragraph{Theoretical convergence.}
\begin{wrapfigure}{r}{5cm}
	\centering
	\includegraphics[width=\linewidth]{patch}
	\caption{The patch around the $i$th sampling point.}
\end{wrapfigure}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling.
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
\begin{equation} \label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=1}^{n} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
\end{equation}
This operator extends the graph Laplacian operator previously defined with the weighting scheme \eqnref{weights:belkin} to each point of the the sphere.
% \mdeff{Can we avoid to introduce general manifolds and stay on the sphere?} \nati{yes}
As the radius of the kernel $t$ will be adapted to the number of samples, we also rescale the graph Laplacian operator as $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$.
Hence, we have $\b{L}_n^t \b{f} = T_\V L_n^t f$.
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling $\V$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
Define $d^{(n)} := \max_{i=1, \dots, n} d_i$ and $A^{(n)} := \max_{i=1, \dots, n} A_i$.
%Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\Rg$ such that for any function $f$ on the sphere
%\mdeff{Already defined in problem formulation.}
%\begin{equation} \label{eq:rotation operator}
%	\Rg f(x) := f(g^{-1}(x)).
%\end{equation}
\begin{theorem}
	For a sampling $\V$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*}
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the rescaled graph Laplacian (the discrete counterpart of $\hat L^t_n$) converges pointwise towards the Laplace-Beltrami operator $\Delta_{\S^2}$ in the sense explained in \citep{belkin2007convergence}.
This is a major step towards equivariance as the Laplace-Beltrami operator does commute with rotation. Based on this property, we show that the graph Laplacian is equivariant to rotation in the limit of infinite sampling.

\begin{theorem}\label{theo:equivariance}
Under the hypothesis of theorem~\ref{theo:pointwise convergence for a regular sampling}, the rescaled graph Laplacian commutes with any rotation, in the limit of infinite sampling, i.e.,
\begin{equation*}
	\forall y\in\mathbb S^2\quad\left| \Rg \hat L_n^{t_n} f (y) - \hat L_n^{t_n} \Rg f(y) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{theorem}

@@

\section{Proof of Theorem~\ref{theo:equivariance}}
\begin{proof}
Fix $x\in \mathbb S^2$. Since any rotation $\Rg$ is an isometry, and the Laplacian $\Delta$ commutes with all isometries of a Riemanniann manifold, and defining $\Rg f =: f'$ for ease of notation, we can write that
	\begin{align*}
			\seminorm{\Rg \Ln f (x) - \Ln \Rg f(x) } &\leq \seminorm{\Rg \Ln f (x) - \Rg \laplacian f(x)} + \seminorm{\Rg \laplacian f (x) - \Ln \Rg f(x)} =\\
			& =\seminorm{\Rg (\Ln f - \laplacian f ) (x)} + \seminorm{ \laplacian f' (x) - \Ln f'(x)} \leq \\
			& \leq \seminorm{(\Ln f - \laplacian f ) (g^{-1}(x))} + \seminorm{ \laplacian f' (x) - \Ln f'(x)}
	\end{align*}
